Ricci-Ollivier Curvature of the Rooted Phylogenetic Subtree-Prune-Regraft Graph\tnoteref{t1,t2}

### 2016-12-01 response_to_reviewers.txt

We thank the reviewers for their suggestions.

We are especially grateful to Reviewer 2, who has offered the most detailed and constructive review we may have ever received.

@@

Reviewer 1

I think that the abstract is too long and should be three-time shorter.
* TODO, any comments Erick?

There are several occurrence of the word “taxon” or “taxa” while there is no definition: I guessed it was the leaves, and maybe the authors want to replace every occurrence.
* replaced each occurrence of taxon or taxa with leaf or leaves.

@@

23: What is the unit of the 'O(n) space'?
* Bits of memory, as is standard for the Big-O notation with respect to space.

52-55: The significance of this paragraph could be increased by stating the special features, properties or outcomes of these runs instead of barely stating that you have performed them.
* The runs are analyzed in the next section. To avoid confusion we moved
	the description of the runs to the Access time simulation section and
	simply say "We used this procedure to sample random walks on varying
	size rSPR graphs for our analysis in the next section" here.

Page 13
32: Which graph is 'the rSPR graph ... as discussed in Section 3.1'? The complete one with (2n-3)!!  trees? This is not quite clear since section 3.1 also considers other values for m.
* This refers to the complete graph. We revised our use of m as
	described above, to limit this type of confusion. In addition we added
	the word full ("full rSPR graph") to this sentence.

33: It is stated that the curvature is computed for 'given pairs of trees', but who / what gives these pairs? The wording is a bit odd here.
* As we now clarify with revised wording, this is explained in the next paragraph. We could not compute
	curvatures for all pairs of trees directly, but instead compute
	curvatures for each tanglegram. The procedure described here is to
	used to compute the curvature of any given pair of trees.

### 2016-12-01 curvature.tex
A few fixes.

\subsection{Computing curvature values.}
\label{sec:results:curvature}
To compute curvature values, we first used \texttt{dense\_spr\_graph} to compute the full rSPR graph for four to seven leaves, as discussed in Section~\ref{sec:computing_treespace}.
We then computed curvatures for given pairs of trees directly, by using linear programming~\citep{Loisel2014-hu} to compute the minimal mass transport $W_1$ using the SAGE \citep{SAGE} front-end to the GLPK~\citep{glpk} solver; code can be found in \citep{gricci} which grew from the code described in \citep{Loisel2014-hu}.

This would have required an enormous amount of computation to directly compute curvatures for the $((2n-3)!!)^2$ pairs of trees with $n$ leaves, even for the small values of $n$ we consider here.
We instead exploited the fact that pairs of trees which are equivalent modulo label renumbering are symmetric in the rSPR graph and therefore guaranteed to have the same curvature.
For example, the pairs
$\{(((1,2),3),4), ((1,2),(3,4))\}$ and
$\{(((1,4),2),3), ((1,4),(2,3))\}$
are the same after relabeling, so their curvatures are the same.
We thus directly computed curvature values for one representative pair from each such equivalence class, or \emph{tanglegram} \citep{Venkatachalam2010-zh}; the group-theoretic enumeration methods are described in \citep{Matsen2015-yv} with a closed formula in \citep{billey2016trees}, and the SAGE \citep{SAGE} and GAP4 \citep{GAP4} code is at \citep{tangle}.

>>>>

\subsection{Computing curvature values.}
\label{sec:results:curvature}
To compute curvature values, we first used \texttt{dense\_spr\_graph} to compute the full rSPR graph for four to seven leaves, as discussed in Section~\ref{sec:computing_treespace}.
We then computed curvatures for pairs of trees directly, by using linear programming~\citep{Loisel2014-hu} to compute the minimal mass transport $W_1$ using the SAGE \citep{SAGE} front-end to the GLPK~\citep{glpk} solver; code can be found in \citep{gricci} which grew from the code described in \citep{Loisel2014-hu}.
Specifically, we did this calculation for all representative pairs of trees as described next.

To directly compute curvatures for all $((2n-3)!!)^2$ pairs of trees with $n$ leaves would have required an enormous amount of computation, even for the small values of $n$ we consider here.
We instead exploited the fact that pairs of trees which are equivalent modulo label renumbering are symmetric in the rSPR graph and therefore guaranteed to have the same curvature.
For example, the pairs
$\{(((1,2),3),4), ((1,2),(3,4))\}$ and
$\{(((1,4),2),3), ((1,4),(2,3))\}$
are the same after relabeling, so their curvatures are the same.
We thus directly computed curvature values for one representative pair from each such equivalence class, or \emph{tanglegram} \citep{Venkatachalam2010-zh}; the group-theoretic enumeration methods are described in \citep{Matsen2016-cv} with a closed formula in \citep{Billey2017-ci}, and the SAGE \citep{SAGE} and GAP4 \citep{GAP4} code is at \citep{tangle}.
